---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Let's Load the Activity data

```{r}
activity <- unzip("activity.zip")
act1 <- read.csv("activity.csv")
act1$date<- as.Date(act1$date,"%Y-%m-%d") #Changes date column to date class
head(act1)
```



## What is mean total number of steps taken per day?


Let's create the total steps per day 

```{r}
pow<-aggregate(act1$steps~act1$date,act1, sum) #totalsteps using aggregate(dataframe)
names(pow)<- c("date","steps")    #renaming columns
head(pow)
```

Let's plot the histogram

```{r}
library(ggplot2)
g <- ggplot(pow,aes(x=date,y=steps))
g1<- g + geom_bar(stat="identity")+ylab("Total Steps Per Day")
g1
```


```{r}
mean(pow$steps)#MEAN total number of steps taken
median(pow$steps)#MEDIAN total number of steps taken
```





## What is the average daily activity pattern?


Let's create the average number of steps taken for each 5-minute interval
and plot the time series

```{r}
pow2<-aggregate(act1$steps~act1$interval,act1, mean)#Average daily activity                                                            pattern
                                                     
names(pow2)<- c("interval","steps")
plot(pow2$interval,pow2$steps, type="l",xlab="interval",ylab="Ave Steps",main= "Average Daily Activity Graph",at = seq(from=0,
to=max(pow2$interval),by=5))
```


```{r}
goom<- pow2[max(pow2$steps),  ]
goom[1,1] # is the 5-minute interval that may contain the highest #of steps
```





## Imputing missing values


Total number of missing values is

```{r}
colSums(is.na(act1))
```

Before we replace missing values look at original data

```{r}
head(act1)
```


NAs replaced with mean number of steps for the respective 5-minute interval:

```{r}
pow3<-aggregate(act1$steps~act1$interval,act1, mean)
 names(pow3)<- c("intervals","steps")  #Means for each interval



act2<- act1#in order not to messup original data act1
 nas<-is.na(act2$steps)
 act2$steps[nas]<- pow3$steps[nas] #Replacing NAs with means of respective 5-minute interval
                                    

head(act2)  ##Original data filled in by mean of respective 5-minute interval
```


MEAN and MEDIAN of total number of steps taken each day are

```{r}
pow4<-aggregate(act2$steps~act2$date,act2, sum)#tsteps using aggregate(dataframe)
names(pow4)<- c("date","steps")#renaming columns

#check with mean and median above. They seem to bethe same
 mean(pow4$steps)

 median(pow4$steps)
```



Data looks like this:

```{r}
 head(pow4)
```


Histogram looks like this:

```{r}
 g <- ggplot(pow4,aes(x=date,y=steps))
g1<- g + geom_bar(stat="identity")+ylab("Total Steps Per Day")
g1
```

Plots look similar in addition to almost identical means and medians.
This shows that replacing missing values with me does not change the results
that much if it affects it at all




## Are there differences in activity patterns between weekdays and weekends?


The following code replaces dates with "Weekday" and "Weekend" and 
then displays a few rows:

```{r}
act22<-act2 #In order not to mess up act2
act22$date<-weekdays(act22$date,abbreviate=F)#date changed to weekdays
head(act22)
```


Now the "Weekdays" and "Weekends" showing

```{r}
happy<- c("Saturday","Sunday")#Weekend vector

 weeke<- act22$date==happy
 act22$date[!weeke]<- "Weekday" ##Replacing week days with "Weekdays"
act22$date[weeke]<- "Weekend"  ##Replacing weekend days with "Weekend"

> head(act22)
```


Let's have a look at the time series plot

```{r}
pow5<-aggregate(act22$steps~act22$date+act22$interval,act22, mean)
 names(pow5)<- c("date","interval","steps")

qplot(interval,steps, data=pow5, ylab="Number of Steps",
     geom = "line", facets= date~.)#time series graph showing weekdays and                                           weekends
                                               
```




